Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance

### 2019-06-29 Manuscript/manuscript.Rmd
edits

In order to set a baseline in performance, we first evaluated the least granular random forest classifier, that is one implemented using OC predictor types, without data re-sampling. We found that models built on data with fully imbalanced classes exhibited poor MCCs across all resolutions (Figure 4A). We gradually increased the complexity in data level characteristics by assessing models built on OP predictor types, while still not performing any data resampling. Here, the results were similar to that of OC types (Figure 4B). However, when including distance-type predictors, models were found to exhibit a marked improvement in MCC compared to the other two types (Figure 4C). Results were found to be consistently poor across resolutions and predictor types, when implemented on the K562 cell line (Supplementary Figure 3). Furthermore, we found that conventional metrics of performance (i.e. accuracy and AUC) were less sensitive to models built on imbalanced data, regardless of cell line, resolution, or predictor-type (Supplementary Tables 6-7).

@@

![**Figure 4. SMOTE re-sampled models built on distance-type predictors yielded optimal MCCs across resolutions for GM12878.** Performances of random forest classification models built on each of the 3 predictor types: (A) OC, (B) OP, and (C) Distance. Within each predictor type, and across resolution, training data was re-sampled using different techniques including: None (red), ROS (green), RUS (blue), and SMOTE (purple). Performances were aggregated using means across each chromosome, with error bars representing 1 standard deviation away from the mean.](Figures/figure4_gm12878_model_performances.png)

>>>>

We first evaluated the least granular random forest classifier, that was one implemented using OC predictor types, without data re-sampling. We found that these models exhibited poor MCCs across all resolutions (Figure 4A (red)). We then assessing models built on OP predictor types, while still not performing any data resampling. Performances were similar to that of OC types, with no clear improvement (Figure 4B (red)). This indicated that the added complexity of using overlap percents from the functional genomic annotations did not affect predictive performance.

However, when implemented using distance-type predictors and no re-sampling, models were found to exhibit a marked improvement in MCC compared to the other two types (Figure 4C (red)).

<!--Add K562 results in main paper-->

<!--
Results were found to be consistently poor across resolutions and predictor types, when implemented on the K562 cell line (Supplementary Figure 3). Furthermore, we found that conventional metrics of performance (i.e. accuracy and AUC) were less sensitive to models built on imbalanced data, regardless of cell line, resolution, or predictor-type (Supplementary Tables 6-7). 
-->

@@

![**SMOTE re-sampled models built on distance-type predictors yielded optimal MCCs across resolutions for GM12878.** Performances of random forest classification models built on each of the 3 predictor types: (A) OC, (B) OP, and (C) Distance. Within each predictor type, and across resolution, training data was re-sampled using different techniques including: None (red), ROS (green), RUS (blue), and SMOTE (purple). Performances were aggregated using means across each chromosome, with error bars representing 1 standard deviation away from the mean.](Figures/figure4_gm12878_model_performances.png)
